Subtle Gestures with the J!NS MEME

### 2021-08-17 Paper/paper.tex
Write introduction for Analysis section Also add an illustration for our workflow and fix some minor mistakes.

% Assigned to FB
\section{Analysis}
% TODO introduction for analysis section
What does our research contribute? (Additional subtle gestures, ...)

% TODO, maybe do this in the Introduction section? ~FB
Explaining the J!ns Meme.

\subsection{Workflow}
We began with planning the rough outline that our research would follow.
The plan was as follows:
\begin{itemize}
    \item Gather gesture ideas
    \item Check ideas for detectable responses
    \item Create tools to automate the recording of gestures
    \item Collect labelled data with the tools
    \item Model, train and validate a neural network until its performance is sufficient
    \item Test the network on our withheld test dataset
\end{itemize}

[Insert figure here]

@@

\begin{table}
\centering
\caption{The quality of the EOG and IMU sensor responses for our gestures. Gestures 
        with (L/R) at the end have both a left and a right variant.}\label{tab:responses}
\begin{tabular}{|l|l|}
    \hline
    Gesture description & Response quality\\
    \hline
    Pushing glasses up the nose & strong\\
    Readjusting glasses (L/R)& strong\\
    Thinking pose & weak\\
    Slow nodding & strong\\
    Tilting head sideways (L/R)& strong\\
    Stroking mustache & weak\\
    Biting lip & weak\\
    Tapping on the nose (L/R)& strong\\
    Rubbing the nose & strong\\
    Pushing cheek up with a finger (L/R)& strong\\
    Tapping the glasses (L/R)& strong\\\hline
\end{tabular}
\end{table}

\subsection{Tools for Recording and Labelling}
% TODO include the links via BibTeX
To use the J!ns Meme, we first had to interface with it. The interface would have to allow
both recording and streaming the data from the smartglasses, perferrably.
The J!ns Meme's sensor data can be read either via an iOS, Android or macOS app
or via a Windows app, with the latter requiring a custom bluetooth dongle. For
initial gesture evaluation we used the Android app, but since we set out to have
live classification with the potential to control your computer, and only Windows
machines were available, we chose the Windows app J!ns Meme DataLogger
\cite{jinsMemeDataLoggerWin}. A software development kit is available for Android, macOS
and Windows as well \cite{jinsMemeSDK}, but it did not eliminate the
need for the bluetooth dongle on Windows and only implied a bigger overhead for us,
since the DataLogger can both record and forward the J!ns Meme's sensor data to a
file and a TCP port, respectively.

%TODO check correctness, this is from memory alone
The J!ns Meme DataLogger supplies eleven values per cycle, three of which are the
accelerometers' data and the remaining eight being processed EOG data.
Notably, we observed that we had different values available via the Android app
\cite{jinsMemeDataLoggerAndroid}. Namely there were gyroscope values, albeit they
were constantly zero, and less EOG data fields.
We found this behavior to be the same on all J!ns Meme smartglasses and Android phones
we had available, but are not certain if it is intended behavior.

>>>>

% Assigned to FB
\section{Analysis}
While there is quite a bit of research about the J!ns Meme and its applications, there
seemed to be room for more gestures to be scientifically explored, especially for the
purpose of subtle interaction. Therefore we set out to experiment with various
gestures to find out how well the J!ns Meme captured them. Our method of detection and
classification was an artificial neural network (ANN). ANNs have proven themselves as
a reliable way of classifying data, be it static data like images or text or sequential
data like audio recordings or sensor readouts.

\subsection{Workflow}
\begin{figure}
\centering
\centerline{\includegraphics[scale=0.8]{workflow.pdf}}
\caption{The workflow of this project}
\label{fig:workflow}
\end{figure}

We began with planning the rough outline that our research would follow.
The plan was as seen in Figure \ref{fig:workflow}:
\begin{itemize}
    \item Gather gesture ideas
    \item Check ideas for detectable responses
    \item Create tools to automate the recording of gestures
    \item Collect labelled data with the tools
    \item Model, train and validate a neural network until its performance is sufficient
    \item Test the network on our withheld test dataset
\end{itemize}

@@

\begin{table}
\centering
\caption{The quality of the EOG and IMU sensor responses for our gestures. Gestures 
        with (L/R) at the end have both a left and a right variant.}\label{tab:responses}
\begin{tabular}{|l|l|}
    \hline
    Gesture description & Response quality\\
    \hline
    Pushing glasses up the nose & strong\\
    Readjusting glasses (L/R) & strong\\
    Thinking pose & weak\\
    Slow nodding & strong\\
    Tilting head sideways (L/R) & strong\\
    Stroking mustache & weak\\
    Biting lip & weak\\
    Tapping on the nose (L/R)& strong\\
    Rubbing the nose & strong\\
    Pushing cheek up with a finger (L/R) & strong\\
    Tapping the glasses (L/R) & strong\\\hline
\end{tabular}
\end{table}

\subsection{Tools for Recording and Labelling}
To use the J!ns Meme, we first had to interface with it. The interface would have to allow
both recording and streaming the data from the smartglasses, perferrably.
The J!ns Meme's sensor data can be read either via an iOS, Android or macOS app
or via a Windows app, with the latter requiring a custom bluetooth dongle. For
initial gesture evaluation we used the Android app, but since we set out to have
live classification with the potential to control your computer, and only Windows
machines were available, we chose the Windows app J!ns Meme DataLogger
\cite{jinsMemeDataLoggerWin}. A software development kit is available for Android, macOS
and Windows as well \cite{jinsMemeSDK}, but it did not eliminate the
need for the bluetooth dongle on Windows and only implied a bigger overhead for us,
since the DataLogger can both record and forward the J!ns Meme's sensor data to a
file and a TCP port, respectively.

%TODO check correctness, this is from memory alone
The J!ns Meme DataLogger supplies eleven values per cycle, three of which are the
accelerometers' data and the remaining eight being at least partially processed EOG data.
Notably, we observed that we had different values available via the Android app
\cite{jinsMemeDataLoggerAndroid}. Namely there were gyroscope values, albeit they
were constantly zero, and less EOG values per time step.
We found this behavior to be the same on all J!ns Meme smartglasses and Android phones
we had available, but are not certain if it is intended behavior.
